---
title: "Circle Packing aRt"
author: "Daniel E. Weeks"
date: "`r format(Sys.time(), '%B %d, %Y, %R')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 3
  html_document:
    df_print: paged
    toc: true
    number_sections: true
    toc_depth: '3'
    code_folding: show
#   md_extensions: +raw_attribute
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r,echo=FALSE}
# In the event of a crash, save the working space on exit:
save_all <- function() {
  save.image("recover.RData")
}
options(error = save_all)
```





# Source

From

https://github.com/will-r-chase/aRt/blob/master/January_tessellated_menagerie/circlepack_image.R

```
## -------
## barely modified from
## https://chichacha.netlify.com/2018/12/22/bubble-packed-chart-with-r-using-packcircles-package/
```

# Load Libraries

```{r load_libraries, message=FALSE}
library(tidyverse) ## I need tidyverse for everything :)
library(imager) ## to create data frame from image
library(scales) ## rescale function is so handy!
library(packcircles) ## making circle packing easy!
```

# Step 1: Load image and convert to data frame

```{r}
im <- load.image("CommonBuckeye_selected.jpg")
plot(im)
```

```{r}
## Convert Image into Data Frame
im.df.colour <- im %>%
  as.data.frame(wide="c") %>% ## so that rgb value is in separate column.
  rename(im_x=x,im_y=y) %>%
  mutate(hex=rgb(c.1,c.2,c.3))
```

# Steps 2 and 3 using circleProgressiveLayout function.

This works best if the source image is square.  

```{r}
## Generate circle packing layout using rbeta distribution as size of circles
pack_layout <- circleProgressiveLayout(rbeta(5000,1,2), sizetype='area') %>%
  ## Step 3 - I want to figure out what colour to use, so I want layout & image df to have same scaling.
  mutate(im_x=floor(rescale(x,to=range(im.df.colour$im_x))),
         im_y=floor(rescale(y,to=range(im.df.colour$im_y))),
         ## also generate id, so i can join the data frame easily later!
         id=row_number()) %>%
  inner_join(im.df.colour %>% select(im_x,im_y,hex), by=c("im_x","im_y"))

#for smaller circles inside
pack_layout2 <-
  pack_layout %>%
  mutate(r = sqrt(x^2 + y^2), ## calculate distance from 0,0 coordinate
         angle_t = atan2(y,x),  ## The arc-tangent of two arguments atan2(y, x) returns the angle between the x-axis and the vector from the origin to (x, y)
         angle = rescale(angle_t, from=c(-pi,pi))) ## convert theta value to value between 0 and 1
```

# Step 4

circleLayoutVertices takes as input a matrix or data.frame of circle data (x, y, size).

```{r}
## Using the layout above create data frame using circleLayoutVertices function so that you can plot circle using ggplot2
data_gg <- circleLayoutVertices(pack_layout) %>%
  inner_join(pack_layout %>% select(id,hex), by=c("id"))

#for smaller circles inside
data_gg2 <- circleLayoutVertices(pack_layout2) %>%
  inner_join(pack_layout2 %>% select(id,hex), by=c("id"))

data_gg2_1 <- circleLayoutVertices(pack_layout2 %>%
                                     sample_n(800, weight=radius) %>%
                                     mutate(radius=0.7*radius), npoints = 25) ## I want to draw smaller circle, so shrink the radius
data_gg2_2 <- circleLayoutVertices(pack_layout2 %>%
                                     sample_n(700,weight=radius) %>%
                                     mutate(radius=0.5*radius), npoints = 25)
data_gg2_3 <- circleLayoutVertices(pack_layout2 %>%
                                     sample_n(900,weight=radius) %>%
                                     mutate(radius=0.3*radius), npoints = 25)
```

# Unscaled plotting on the original scales

To adjust for the fact that the original source image is not square, here instead we use the `pack_layout` on the original scale, but then we have to adjust the `radius`.  The adjustment was determined by repeated plotting with different levels of adjustment.

```{r}
pack_layout3 <- pack_layout %>% select(im_x,im_y,radius,id,hex) %>% 
  mutate(x = im_x, y= im_y, radius=22*radius) %>% 
  select(x,y,radius,id,hex)
dim(pack_layout3)
head(pack_layout3)
data_gg3 <- circleLayoutVertices(pack_layout3) %>%
  inner_join(pack_layout3 %>% select(id,hex), by=c("id"))
```




```{r}
## plot bubbles
data_gg3 %>%
  ggplot(aes(x=x,y=y,group=id)) +
  geom_polygon(aes(fill=hex)) +
  scale_fill_identity() +
  coord_equal() +
  scale_y_reverse() +  ## you need to reverse y-axis
  theme_void()
```

```{r}
ggsave("CommonBuckeyeCircles_unscaled1.png", device = "png", width = 7, height = 5, type = "cairo")
```


# Plot 1: bubbles

```{r}
## plot bubbles
data_gg %>%
  ggplot(aes(x=x,y=y,group=id)) +
  geom_polygon(aes(fill=hex)) +
  scale_fill_identity() +
  coord_equal() +
  scale_y_reverse() +  ## you need to reverse y-axis
  theme_void()
```

Original image is 1455 × 987

```{r}
ggsave("CommonBuckeyeCircles1.png", device = "png", width = 5, height = 5, type = "cairo")
```

# Plot 2: with smaller circles inside

```{r}
#with smaller circles inside
data_gg2 %>%
  ggplot(aes(x=x,y=y, group=id)) +
  geom_polygon(aes(fill=hex)) +
  geom_path(data=data_gg2, size=0.5, color="#ffffff90") +
  geom_path(data=data_gg2_1,size=1, color="#ffffff90") +
  geom_path(data=data_gg2_2,size=0.5, color="#ffffff90") +
  geom_path(data=data_gg2_3,size=0.3, color="#ffffff90") +
  scale_fill_identity() +
  scale_color_identity() +
  theme_void() +
  coord_fixed() +
  scale_y_reverse()

ggsave("CommonBuckeyeCircles2.png", device = "png", width = 5, height = 5, type = "cairo")

saveRDS(data_gg, "CommonBuckeye_points.rds")
```


# Unscaled Plot 2 

Here we redraw the version with the smaller circles on the inside on the original scale, to compensate for the fact that the original source image is not square.


```{r}

#for smaller circles inside
pack_layout4 <-
  pack_layout3 %>%
  mutate(r = sqrt(x^2 + y^2), ## calculate distance from 0,0 coordinate
         angle_t = atan2(y,x),  ## The arc-tangent of two arguments atan2(y, x) returns the angle between the x-axis and the vector from the origin to (x, y)
         angle = rescale(angle_t, from=c(-pi,pi))) ## convert theta value to value between 0 and 1

data_gg2 <- circleLayoutVertices(pack_layout4) %>%
  inner_join(pack_layout3 %>% select(id,hex), by=c("id"))

data_gg2_1 <- circleLayoutVertices(pack_layout4 %>%
                                     sample_n(800, weight=radius) %>%
                                     mutate(radius=0.7*radius), npoints = 25) ## I want to draw smaller circle, so shrink the radius
data_gg2_2 <- circleLayoutVertices(pack_layout4 %>%
                                     sample_n(700,weight=radius) %>%
                                     mutate(radius=0.5*radius), npoints = 25)
data_gg2_3 <- circleLayoutVertices(pack_layout4 %>%
                                     sample_n(900,weight=radius) %>%
                                     mutate(radius=0.3*radius), npoints = 25)
```


```{r}
#with smaller circles inside
data_gg2 %>%
  ggplot(aes(x=x,y=y, group=id)) +
  geom_polygon(aes(fill=hex)) +
  geom_path(data=data_gg2, size=0.5, color="#ffffff90") +
  geom_path(data=data_gg2_1,size=1, color="#ffffff90") +
  geom_path(data=data_gg2_2,size=0.5, color="#ffffff90") +
  geom_path(data=data_gg2_3,size=0.3, color="#ffffff90") +
  scale_fill_identity() +
  scale_color_identity() +
  theme_void() +
  coord_fixed() +
  scale_y_reverse()
```

```{r}
ggsave("CommonBuckeyeCircles_unscaled2.png", device = "png", width = 7, height = 5, type = "cairo")
```

# Session Information

```{r}
sessionInfo()
```

