---
title: "Heart Art"
author: "Daniel E. Weeks"
date: "`r format(Sys.time(), '%B %d, %Y, %R')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: true
    toc_depth: '3'
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 3
#   md_extensions: +raw_attribute
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r,echo=FALSE}
# In the event of a crash, save the working space on exit:
save_all <- function() {
  save.image("recover.RData")
}
options(error = save_all)
```


# Load Libraries

```{r load_libraries,message=FALSE}
library(tidyverse)
# library(tidylog)
library(patchwork)
library(pryr)
```

# Input directory and files 

```{r}
# Print the working directory
getwd()
```
The input files used in this analysis are ... and can be found in ... 

# Heart Art

The goal here was to create some art out of hearts for use in making a Valentine's Day card.

## Polar coordinate heart 1

From

https://stackoverflow.com/questions/8082429/plot-a-heart-in-r

```{r}
dat <- data.frame(x = seq(0, 2 * pi, length.out = 100))
cardioid <- function(x, a = 2) {
  a * (1 - cos(x))
}
ggplot(dat, aes(x = x)) + stat_function(fun = cardioid) + coord_polar()
```
## Polar coordinate heart 2

```{r}
x = seq(0, 2 * pi, length.out = 100)
heart <- function(x) {
  2 - 2 * sin(x) + sin(x) * (sqrt(abs(cos(x)))) / (sin(x) + 1.4)
}
ggplot(dat, aes(x = x)) + stat_function(fun = heart) + coord_polar(start =
                                                                     -pi / 2)
```
# X, Y heart

From https://rpubs.com/RandhyllCho/250112

```{r}
dat <- data.frame(t = seq(0, 2 * pi, by = 0.01))
x <-  function(t)
  16 * sin(t) ^ 3
y <- function(t)
  13 * cos(t) - 5 * cos(2 * t) - 2 * cos(3 * t) - cos(4 * t)

dat$y <- y(dat$t)
dat$x <- x(dat$t)

heart <- ggplot(dat, aes(x, y)) +
  geom_polygon(fill = "red", col = "firebrick", alpha = 0.9) +
  theme_classic()

heart
```

# X, Y heart 2

From https://www.r-bloggers.com/2019/01/making-original-bingo-heart-theme/

```{r}
# library(tidyverse)
# library(patchwork)

##  Function to generate heart shape around point xc and rc with some sizing.  Output of function is data frame with bunch of points which you can use to draw a heart!
# http://mathworld.wolfram.com/HeartCurve.html
heart_vertices <- function(xc, yc, size, npoints = 100, ...) {
  #area = pi*r*r for circle... this heart has area of 180 = r*r
  radius = size * 0.05  ## I'm not set on this...  I just wanted to make size smaller
  yc = yc + 0.1 * radius  ## Just adjusting center of heart bit
  t = seq(-pi, pi, length.out = npoints + 1)
  x = xc + 16 * radius * (sin(t)) ^ 3
  y = yc + radius * 13 * cos(t) - radius * 5 * cos(2 * t) - radius * 2 *
    cos(3 * t) - radius * cos(4 * t)
  df <- tibble(theta = t, x = x, y = y)
  return(df)
}

## Above function with generate points you'd need to generate heart. If you adjust npoints to be higher, then you can draw smoother shape.

heart_vertices(0,0,1) %>%
  ggplot() +
  geom_polygon(aes(x=x,y=y), fill="pink", 
               color="#61537520",size=10) + 
  theme_void() +
  coord_fixed() 
```

# Grid of hearts

From https://www.r-bloggers.com/2019/01/making-original-bingo-heart-theme/

```{r}
xg <- 8
# yg = # of grid points
# yg is limited by the number of values in the scale_fill_manual
yg <- 5
my_grid <- tibble(
  xc=rep(c(1:xg), times=yg), 
  yc=rep(c(1:yg), each=xg),
  size=0.6, 
  id = c(1:(xg*yg))
) 

# pmap applies the function to each row in the data frame.

## For each point on grid generate points to draw heart 
heart_grid <- my_grid %>%
  mutate(v = pmap(., heart_vertices)) %>%
  unnest(v) %>%
  ggplot(aes(x = x, y = y, group = id)) +
  geom_polygon(aes(fill = factor(yc))) +
  coord_fixed() +
  theme_void() +
  scale_x_continuous(breaks = seq(0.5, xg+0.5, by = 1), labels = NULL) +
  scale_y_continuous(breaks = seq(0.5, yg+0.5, by = 1), labels = NULL) +
  scale_fill_manual(
    values = c("#30bcad", "#57a337", "#f8b620", "#e03426", "#eb73b3"),
    guide = "none"
  )

heart_grid
```

```{r}
ggsave("heart_grid.jpeg", width=4, height=4, units = "in", dpi=300)
```


```{r}
heart_grid <- my_grid %>%
  mutate(v = pmap(., heart_vertices)) %>%
  unnest(v) %>%
  ggplot(aes(x = x, y = y, group = id)) +
  geom_polygon(aes(fill = factor(yc))) +
  coord_fixed() +
  theme_void() +
  scale_x_continuous(breaks = seq(0.5, xg+0.5, by = 1), labels = NULL) +
  scale_y_continuous(breaks = seq(0.5, yg+0.5, by = 1), labels = NULL) +
  scale_fill_manual(
    values = c("deeppink4", "deeppink3", "deeppink2", "deeppink1", "deeppink"),
    guide = "none"
  )

heart_grid
```

```{r}
ggsave("heart_grid_deeppink.jpeg", width=6, height=4, units = "in", dpi=300)
```

```{r}
heart_grid <- my_grid %>%
  mutate(v = pmap(., heart_vertices)) %>%
  unnest(v) %>%
  ggplot(aes(x = x, y = y, group = id)) +
  geom_polygon(aes(fill = factor(yc))) +
  coord_fixed() +
  theme_void() +
  scale_x_continuous(breaks = seq(0.5, xg+0.5, by = 1), labels = NULL) +
  scale_y_continuous(breaks = seq(0.5, yg+0.5, by = 1), labels = NULL) +
  scale_fill_manual(
    values = c("hotpink4", "hotpink3", "hotpink2", "hotpink1", "hotpink"),
    guide = "none"
  )

heart_grid
```


# Broken Heart

```
x	=	sin(t)*cos(t)*log(abs(t))	
y	=	(abs(t)^(0.3))*(cos(t))^(1/2)	
```

where t is in (-1, 1)

```{r}
heart_vertices <- function(xc, yc, size, npoints = 100, ...) {
  t = seq(-1, 1, length.out = npoints + 1)
  x	=	sin(t)*cos(t)*log(abs(t))	
  y	=	abs(t)^(0.3)*sqrt(cos(t))	
  df <- tibble(t = t, x = x, y = y)
  return(df)
}

heart_vertices(0,0,1) %>%
  ggplot() +
  geom_polygon(aes(x=x,y=y), fill="pink", 
               color="#61537520",size=10) +  
  theme_void() +
  coord_fixed() 

```

# Heart Scatter Plot

From: https://www.summitllc.us/blog/show-your-valentine-how-much-you-care-create-a-statistical-graph

```{r}
# heart curve formula
heart <- quote((x^2 + y^2 - 1)^3 - x^2 * y^3)

# formula for heart curve at a given x
heart_at_x <- function(x) {
  function(y) eval(substitute_q(heart, list(x = x)), list(y = y))
}

# trace the heart curve
# by evaluating the heart curve formula at each x, then finding the roots of the
# resulting formula in y; e.g. a x==0, find the roots of (y^2 - 1)^3 = 0
# broken up into upper and lower parts (h_y1 and h_y2)
heart_x <- seq(-1.136, 1.136, 0.001)
heart_y_lower <- sapply(heart_x, function(x) uniroot(heart_at_x(x), c(-2, 0.6))$root)
heart_y_upper <- sapply(heart_x, function(x) uniroot(heart_at_x(x), c(0.6, 2))$root)

# put together data frame
heart_df <- data.frame(x = rep(heart_x, 2), 
                       y = c(heart_y_lower, heart_y_upper))

# show outline
with(heart_df, plot(x, y))

# create a data frame with one row per x, so we can fill in the heart
heart_df_minmax <- data.frame(x = heart_x,  
                              y_min = heart_y_lower, 
                              y_max = heart_y_upper)

# set.seed(20150214)

# fill in the heart by generating random deviates at each x 
# and rejecting those that fall outside the heart curve
heart_full <- apply(heart_df_minmax, 
                    1, 
                    function(w) {
                      x <- w["x"]
                      y_min = w["y_min"]
                      y_max = w["y_max"]
                      y <- rnorm(2, mean = 0.33)
                      y <- y[between(y, y_min, y_max)]
                      x <- x[any(is.finite(y))]
                      data.frame(x, y, row.names = NULL)
                    })

# change from list to data frame
heart_full <- bind_rows(heart_full)

# add random numbers for color and size
heart_full <- heart_full %>% 
  mutate(z1 = runif(n()), 
         z2 = pmin(abs(rnorm(n())), 3), 
         order = runif(n())) %>%
  arrange(order)

#------------------------------------------------------------------------------#

# plot the heart
p <- ggplot(heart_full, 
            aes(x, y, color = z1, size = z2)) + 
  geom_point(pch = -1 * as.hexmode(9829)) + 
  scale_color_gradient(limits = c(0, 1), low = "red3", high = "pink") + 
  scale_size(limits = c(0, 3), range = c(0.1, 20)) + 
  theme_void() +
  theme(legend.position="none") 
p

# Use less of the points
p <- ggplot(heart_full[c(1:round(3*nrow(heart_full)/8,0)),], 
            aes(x, y, color = z1, size = z2)) + 
  geom_point(pch = -1 * as.hexmode(9829)) + 
  scale_color_gradient(limits = c(0, 1), low = "red3", high = "pink") + 
  scale_size(limits = c(0, 3), range = c(0.1, 20)) + 
  theme_void() +
  theme(legend.position="none") 
p
```

```{r}
ggsave("heart_scatterplot.jpeg", width=6, height=4, units = "in", dpi=300)
```


# Session Information

```{r}
sessionInfo()
```

