---
title: "Math Art Fun"
author: "Daniel E. Weeks"
date: "`r format(Sys.time(), '%B %d, %Y, %R')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 3
  html_document:
    df_print: paged
    toc: true
    number_sections: true
    toc_depth: '3'
    code_folding: show
#   md_extensions: +raw_attribute
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r,echo=FALSE}
# In the event of a crash, save the working space on exit:
save_all <- function() {
  save.image("recover.RData")
}
options(error = save_all)
```


# Load Libraries

```{r load_libraries,message=FALSE}
library(mathart)
# library(ggart)
library(ggforce)
library(Rcpp)
library(tidyverse)
```

# Input directory and files 

```{r}
# Print the working directory
getwd()
```
The input files used in this analysis are ... and can be found in ... 

# The mathart R package

These are drawn with functions from the `mathart` R package from

https://github.com/marcusvolz/mathart

# Mollusc

```
df <- mollusc()
df1 <- df %>% mutate(id = 1)
df2 <- df %>% mutate(id = 2)
df3 <- df %>% mutate(id = 3)

p <- ggplot() +
  geom_point(aes(x, y), df1, size = 0.03, alpha = 0.03) +
  geom_path( aes(x, y), df1, size = 0.03, alpha = 0.03) +
  geom_point(aes(x, z), df2, size = 0.03, alpha = 0.03) +
  geom_path( aes(x, z), df2, size = 0.03, alpha = 0.03) +
  geom_point(aes(y, z), df3, size = 0.03, alpha = 0.03) +
  geom_path( aes(y, z), df3, size = 0.03, alpha = 0.03) +
  facet_wrap(~id, nrow = 2, scales = "free") +
  theme_blankcanvas(margin_cm = 0.5)

p
```

```{r}
# Set parameters (see mathart::mollusc() documentation for details)
n_s <- 650L
n_t <- 2000L
n <- 1000
alpha <- 82.6
beta <- 1.515
phi <- 14.3
mu <- 0
Omega <- 0
s_min <- -193.8
s_max <- 69.4
A <- 7.031
a <- 2.377
b <- 6.42
P <- 0
W_1 <- 1
W_2 <- 1
N <- 0
L <- 0
D <- 1
theta_start <- 0
theta_end <- 10*pi

# Generate data
df <- mollusc(n_s = n_s, n_t = n_t,
              alpha = alpha, beta = beta, phi = phi, mu = mu, Omega = Omega, s_min = s_min, s_max = s_max,
              A = A, a = a, b = b, P = P, W_1 = W_1, W_2 = W_2, N = N, L = L, D = D,
              theta_start = theta_start, theta_end = theta_end)

# Create plot
p <- ggplot() +
  geom_point(aes(x, z), df, size = 0.03, alpha = 0.03) +
  geom_path(aes(x, z), df, size = 0.03, alpha = 0.03) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

```{r}
df <- mollusc()

p <- ggplot() +
  geom_point(aes(x, z), df, size = 0.03, alpha = 0.03) +
  geom_path(aes(x, z), df, size = 0.03, alpha = 0.03) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Harmonograph

```{r}
df <- harmonograph(A1 = 1, A2 = 1, A3 = 1, A4 = 1,
                    d1 = 0.004, d2 = 0.0065, d3 = 0.008, d4 = 0.019,
                    f1 = 3.001, f2 = 2, f3 = 3, f4 = 2,
                    p1 = 0, p2 = 0, p3 = pi/2, p4 = 3*pi/2) %>% mutate(id = 1)

p <- ggplot() +
  geom_path(aes(x, y), df, alpha = 0.25, size = 0.5) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## My Harmonograph 1

```{r}
df <- harmonograph(A1 = 1, A2 = 1, A3 = 1, A4 = 1,
                    d1 = 0.006, d2 = 0.0065, d3 = 0.008, d4 = 0.009,
                    f1 = 2.001, f2 = 3, f3 = 3, f4 = 2,
                    p1 = 0, p2 = 0, p3 = pi/4, p4 = 3*pi/4) %>% mutate(id = 1)

p <- ggplot(mapping=aes(x, y), data = df) +
  geom_path(colour="darkblue",alpha = 0.25, size = 1.0) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

```{r}
ggsave("harmonograph01.png", p, width = 20, height = 20, units = "cm", dpi = 300)
```


## My Harmonograph 2

```{r}
df <- harmonograph(A1 = 2, A2 = 1, A3 = 1, A4 = 1,
                    d1 = 0.006, d2 = 0.0065, d3 = 0.008, d4 = 0.009,
                    f1 = 2.001, f2 = 4, f3 = 3, f4 = 2,
                    p1 = pi/4, p2 = pi/4, p3 = pi/4, p4 = 3*pi/16) %>% mutate(id = 1)

p <- ggplot(mapping=aes(x, y), data = df) +
  geom_path(colour="royalblue",alpha = 0.25, size = 1.0) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

```{r}
ggsave("harmonograph02.png", p, width = 20, height = 20, units = "cm", dpi = 300)
```

# Lissajous curves

```{r}
set.seed(1)

df <- 1:20 %>% map_df(~lissajous(a = runif(1, 0, 10), A = runif(1, 0, 1)), .id = "id")

p <- ggplot() +
  geom_path(aes(x, y), df, size = 0.25, lineend = "round") +
  facet_wrap(~id, nrow = 2) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 0.5)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Rose curves

Rose curves defined by $r=cos(k\theta)$, for various values of $k=n/d$, $n$ across, $d$ down.

```{r}
df <- data.frame(x = numeric(0), y = numeric(0), n = integer(0), d = integer(0))

for(n in 1:6) {
  for(d in 1:6) {
    result <- rose_curve(n, d) %>% mutate(n = n, d = d)
    df <- rbind(df, result)
  }
}

p <- ggplot() +
  geom_path(aes(x, y), df, size = 0.35, lineend = "round") +
  facet_grid(d ~ n) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 1)

```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Fractal ferns

```{r}
n <- 25000

params1 <- data.frame(
  a <- c(0, 0.85, 0.2, -0.15),
  b <- c(0, 0.04, -0.26, 0.28),
  c <- c(0, -0.04, 0.23, 0.26),
  d <- c(0.16, 0.85, 0.22, 0.24),
  e <- c(0, 0, 0, 0),
  f <- c(0, 1.6, 1.6, 0.44),
  p <- c(0.01, 0.85, 0.07, 0.07)
)

params2 <- data.frame(
  a <- c(0, 0.85, 0.09, -0.09),
  b <- c(0, 0.02, -0.28, 0.28),
  c <- c(0, -0.02, 0.3, 0.3),
  d <- c(0.25, 0.83, 0.11, 0.09),
  e <- c(0, 0, 0, 0),
  f <- c(-0.14, 1, 0.6, 0.7),
  p <- c(0.02, 0.84, 0.07, 0.07)
)

df1 <- fractal_fern(n = n, a = params1$a, b = params1$b, c_ = params1$c, d = params1$d, e = params1$e,
                   f = params1$f, p = params1$p) %>% mutate(id = 1)

df2 <- fractal_fern(n = n, a = params2$a, b = params2$b, c_ = params2$c, d = params2$d, e = params2$e,
                    f = params2$f, p = params2$p) %>% mutate(id = 2)

df <- rbind(df1, df2 %>% mutate(x = x*1.75, y = y*1.75))

p <- ggplot() +
  geom_point(aes(x, y), df, size = 0.03, alpha = 0.06) +
  coord_equal() +
  facet_wrap(~id, nrow = 1) +
  theme_blankcanvas(margin_cm = 1)
```


```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Rapidly exploring random tree

```{r}
set.seed(5)
tmp <- capture.output(df <- rapidly_exploring_random_tree(n=1000) %>% mutate(id = 1:nrow(.)))

# Create plot
p <- ggplot() +
  geom_segment(aes(x, y, xend = xend, yend = yend, size = -id, alpha = -id), df, lineend = "round") +
  coord_equal() +
  scale_size_continuous(range = c(0.1, 0.75)) +
  scale_alpha_continuous(range = c(0.1, 1)) +
  theme_blankcanvas(margin_cm = 0)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Vertiginous Spiral

https://www.r-bloggers.com/2018/09/tweetable-mathematical-art-with-r/

```{r}
# library(tidyverse)
ID <- 1
df <- data.frame(x=0, y=0 , value="black", id = ID)
for (i in 2:500){
  df[i,1] <- df[i-1,1]+((0.98)^i)*cos(i)
  df[i,2] <- df[i-1,2]+((0.98)^i)*sin(i) 
  df[i,3] <- ifelse(i %% 2 == 0, "white","black")
  df[i,4] <- ifelse(i %% 3 == 1, ID <- ID + 1, ID)
}

col_list<-c("#FF222C", "#1DFBFF", "#FDFF24", "#2CFF18", "#FF38F4", "#C3C4C9", "#000000")
col_list <- c("red", "chocolate4")
col_list2 <- c("deepskyblue2", "peachpuff")


ggplot(df, aes(x,y)) + 
  geom_polygon(aes(col=value))+
  theme_void() +
  theme(legend.position = "none")

ggplot(df, aes(x,y)) + 
  geom_polygon(aes(col=value, group=id, fill=value))  +
  scale_color_manual(values = col_list) +
  scale_fill_manual(values = col_list2) +
  theme_void() +
  theme(legend.position = "none") 

ggplot(df, aes(x,y, col=factor(id), fill=factor(id))) + 
  geom_polygon(aes( group=id))+
  theme_void() +
  theme(legend.position = "none")
```

# Birds

## Stork

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/mathematical-birds_b_8876904.html)

```{r}
# df <- rbind(parrot() %>% mutate(id = 1),
#            stork() %>% mutate(id = 2),
#            magpie() %>% mutate(id = 3))

df <- stork()

p <- ggplot() +
  geom_point(aes(x, y, alpha = r), df, size = 0.03) +
# facet_wrap(~id, nrow = 2, scales = "free") +
  scale_alpha_continuous(range = c(0.03, 0.06)) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Parrot

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/mathematical-birds_b_8876904.html)

```{r}

df <- parrot()

p <- ggplot() +
  geom_point(aes(x, y, alpha = r), df, size = 0.03) +
# facet_wrap(~id, nrow = 2, scales = "free") +
  scale_alpha_continuous(range = c(0.03, 0.06)) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Magpie

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/mathematical-birds_b_8876904.html)

```{r}

df <- magpie()

p <- ggplot() +
  geom_point(aes(x, y, alpha = r), df, size = 0.03) +
# facet_wrap(~id, nrow = 2, scales = "free") +
  scale_alpha_continuous(range = c(0.03, 0.06)) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Bird in Flight

```{r}
df <- bird_in_flight()

p <- ggplot() +
geom_segment(aes(x, y, xend=xend, yend=yend), df, size = 0.03) +
# facet_wrap(~id, nrow = 2, scales = "free") +
scale_alpha_continuous(range = c(0.03, 0.06)) +
theme_blankcanvas(margin_cm = 1)

```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Critters

## Ant

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/ant-spider-and-millipede-_b_8583338.html)


```{r}
df1 <- ant() %>% mutate(id = 1)
df2 <- spider() %>% mutate(id = 2)
df3 <- millipede() %>% mutate(id = 3)

df <- rbind(df1 %>% mutate(x = (x - min(df1$x))*0.9  + 0.2, y = (y - min(df1$y))*0.9 + 1, r = r * 0.75),
            df2 %>% mutate(x = (x - min(df2$x))*1.5  + 2.4, y = (y - min(df2$y))*1.5 + 2.2),
            df3 %>% mutate(x = (x - min(df3$x))*1.75 + 2.4, y = (y - min(df3$y))*1.75))

df1 <- df %>% filter(id==1)
p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df1, size = 0.03, alpha = 0.03, n = 100) +
  geom_path(aes(x, y, group = id), df1, size = 0.5, alpha = 0.5) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Spider

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/ant-spider-and-millipede-_b_8583338.html)


```{r}
df1 <- df %>% filter(id==2)
p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df1, size = 0.03, alpha = 0.03, n = 100) +
  geom_path(aes(x, y, group = id), df1, size = 0.5, alpha = 0.5) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Millipede

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/ant-spider-and-millipede-_b_8583338.html)


```{r}
df1 <- df %>% filter(id==3)
p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df1, size = 0.03, alpha = 0.03, n = 100) +
  geom_path(aes(x, y, group = id), df1, size = 0.5, alpha = 0.5) +
  coord_equal() +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Butterflies

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/using-mathematical-formul_b_9313484.html)

```{r}
df <- rbind(butterfly1() %>% mutate(id = 1),
            butterfly2() %>% mutate(id = 2),
            butterfly3() %>% mutate(id = 3))

p <- ggplot() +
  geom_point(aes(x, y, alpha = r), df, size = 0.03) +
  coord_equal() +
  facet_wrap(~id, nrow = 2) +
  scale_alpha_continuous(range = c(0.03, 0.16)) +
  theme_blankcanvas(margin_cm = 0)

```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Plants

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/mathematical-plants_b_9010790.html)

## Olive branch

```{r}
df <- rbind(olive_branch() %>% mutate(id = 1),
            palm_branch() %>% mutate(id = 2),
            branch() %>% mutate(id = 3))

df <- olive_branch() %>% mutate(id = 1)

p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df, size = 0.03, alpha = 0.1) +
  coord_equal() +
  # facet_wrap(~id, nrow = 3) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Palm branch

```{r}
df <- palm_branch() %>% mutate(id = 2)

p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df, size = 0.03, alpha = 0.1) +
  coord_equal() +
  # facet_wrap(~id, nrow = 3) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```


## Branch

```{r}
df <- branch() %>% mutate(id = 3)

p <- ggplot() +
  geom_circle(aes(x0 = x, y0 = y, r = r), df, size = 0.03, alpha = 0.1) +
  coord_equal() +
  # facet_wrap(~id, nrow = 3) +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Hearts

Refer to [Hamid's equations](https://www.huffingtonpost.com/entry/mathematical-formulas-for-the-symbol-of-love_b_8848264.html)

## Heart 1

```{r}
df <- rbind(heart1() %>% mutate(id = 1),
            heart2() %>% mutate(id = 2),
            heart3() %>% mutate(id = 3),
            heart4() %>% mutate(id = 4))
i<- 1
p <- ggplot() +
  geom_path(aes(x, y), df %>% filter(id==i), size = 0.25, lineend = "round") +
  # facet_wrap(~id, nrow = 2, scales = "free") +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Heart 2

```{r}
i<- 2
p <- ggplot() +
  geom_path(aes(x, y), df %>% filter(id==i), size = 0.25, lineend = "round") +
  # facet_wrap(~id, nrow = 2, scales = "free") +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

## Heart 3

```{r}
i<- 3
p <- ggplot() +
  geom_path(aes(x, y), df %>% filter(id==i), size = 0.25, lineend = "round") +
  # facet_wrap(~id, nrow = 2, scales = "free") +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```


## Heart 4

```{r}
i<- 4
p <- ggplot() +
  geom_path(aes(x, y), df %>% filter(id==i), size = 0.25, lineend = "round") +
  # facet_wrap(~id, nrow = 2, scales = "free") +
  theme_blankcanvas(margin_cm = 1)
```

```{r, dpi=600, dev='png', echo=FALSE}
p
```

# Other approaches

See this blog by Antonio Sánchez Chinchón:

https://fronkonstin.com

https://github.com/aschinchon

# Session Information

```{r}
sessionInfo()
```

